Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-11-10 demystify.tex
updates

    {\fontsize{16}{19.2} \selectfont \textcolor{ForestGreen}{ \bf ThanhVu (Vu) Nguyen}}\\[0.1in]
    George Mason University, Dept. of Computer Science\\[0.1in]
    \today{} (latest version available on  \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github})
    };
    \end{tikzpicture}

@@

% though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU
So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
Originally, this document was intended for \emph{international students}, but I have expanded it to include information that might also be useful for \emph{US domestic students}.
Moreover, while this primarily aims at students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other good R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines. Currently, 146 (out of 4000) US universities are classified as R1.} institutions in the US.

@@

\sectioninfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance. However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so do contact and introduce yourself.}

@@

\section{How to call or address a professor?}\label{sec:address}
\sectioninfo{What should you call your profs.? Many possibilities (e.g., prof., Dr., and even their firstnames), but not Mr. or Mrs.}

@@

\chapter{MS Admission}\label{chap:ms}
\chapterinfo{MS focuses on coursework and prepares you for \emph{industry}, while PhD focuses on research and prepares you for academia or research.}

@@

\paragraph{About Me} I am an associate professor in the CS dept at George Mason University (Mason). Before Mason, I was at the University of Nebraska-Lincoln (UNL). I have been in the PhD admission process at Mason and UNL for many years. This allows me to have a good understanding of the process, the challenges students face, and what faculty are looking for. I also serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD as discussed in \autoref{chap:ms}). My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.

@@

\textbf{Thank you!}

>>>>

    {\fontsize{16}{19.2} \selectfont \textcolor{ForestGreen}{\bf ThanhVu (Vu) Nguyen}}\\[0.1in]
    George Mason University, Dept. of Computer Science\\[0.1in]
    \today{} (latest version available on  \href{https://github.com/nguyenthanhvuh/phd-cs-us}{Github})
    };
    \end{tikzpicture}

@@

% though \emph{very} top schools could be very selective, e.g., see the \href{https://da-data.blogspot.com/2015/03/reflecting-on-cs-graduate-admissions.html}{admission process} at CMU
So I want to share some details about the admission process and advice for those who are interested in applying for a \textbf{PhD in Computer Science in the US}.
Originally, this document was intended for \emph{international students}, but has been expanded to include information for \emph{US domestic students} (\autoref{sec:domestic-students}).
Moreover, while this primarily aims at students interested in CS, it might be relevant to students from various STEM (Science, Technologies, Engineering, and Mathematics) disciplines.
Furthermore, although many examples are specifics for schools that I and other contributors of this document know about, the information should be generalizable to other good R1\footnote{An \href{https://en.wikipedia.org/wiki/List_of_research_universities_in_the_United_States}{R1 institution} in the US is a research-intensive university with a high level of research activity across various disciplines. Currently, 146 (out of 4000) US universities are classified as R1.} institutions in the US.

@@

\sectioninfo{Individually faculty member \emph{cannot directly admit} a student---so do not email and ask if you have a chance. However, faculty can \emph{advocate} for a student and therefore increase their admission chance---so contact and introduce yourself.}

@@

\section{How to call or address a professor?}\label{sec:address}
\sectioninfo{What should you call your profs.? Many possibilities (e.g., prof., Dr., and even their firstnames), but not Mr., Mrs., or their firstname in the first email.}

@@

\chapter{MS Admission}\label{chap:ms}
\chapterinfo{MS focuses on coursework and prepares you for \emph{industry}, while PhD focuses on research and prepares you for academia or research. So admission requirements are different.}

@@

\paragraph{About Me} I am an associate professor in the CS dept at George Mason University (Mason). Before Mason, I was at the University of Nebraska-Lincoln (UNL). Both UNL and Mason are large R1 research universities in the US and have many international students in their CS graduate programs.
I have been in the PhD admission process at Mason and UNL for many years. This allows me to have a good understanding of the process, the challenges students face, and what faculty are looking for. Currently I serve as the program director of the MS program in Software Engineering at Mason (thus also have experience with MS admission---which is quite different than PhD as discussed in \autoref{chap:ms}). My personal and lab website is at \href{https://dynaroars.github.io}{dynaroars.github.io}.

@@
